\newpage
\section{Verification Failure - Bug Injection}

To demonstrate the robustness of the testbench, specifically the procedural watchdog and concurrent assertions, I forced a bug into the DUT to simulate a designer error.

\subsection{Bug Description: Premature IDLE Transition}
In the correct implementation, when the iterative subtraction loop inside the \texttt{RUN} state achieves convergence (\texttt{a\_next == b\_next}), the FSM must register the result and transition to the \texttt{DONE} state to initiate the output handshake. 

For this failure scenario, the RTL was modified to simulate a designer error where the FSM bypasses the \texttt{DONE} state and transitions directly back to \texttt{IDLE}. Consequently, the \texttt{out\_valid} signal is never asserted for the consumer. 

The injected bug in \texttt{gcd.sv} is shown below:

\begin{minted}[frame=single,framesep=10pt,breaklines]{SystemVerilog}
                // ----------------------------------------------------------
                // RUN: subtractive Euclidean algorithm
                // ----------------------------------------------------------
                RUN: begin
                    a_reg <= a_next;
                    b_reg <= b_next;

                    // Check convergence
                    if (a_next == b_next) begin
                        result_reg <= a_next;
                        // Forced BUG: Should be state <= DONE;
                        state      <= IDLE; 
                    end
                end
\end{minted}

\subsection{Testbench Detection and Logs}
Because the DUT skips the \texttt{DONE} state, the output handshake logic of the uvm driver becomes permanently stalled waiting for \texttt{out\_valid == 1}. This would cause an infinite loop.

However, the injected bug is immediately caught by two distinct mechanisms in the \texttt{gcd\_if} interface:
\begin{itemize}
    \item the procedural watchdog calculates the exact maximum cycles required for the specific operands. When the DUT fails to assert \texttt{out\_valid} within this window, it throws a \texttt{\$fatal} error.
    \item the SVA \texttt{assert\_no\_infinite\_hang} independently monitors the theoretical global timeout limit and throws an error when the transaction exceeds the maximum theoretical latency.
\end{itemize}

Running the \texttt{gcd\_random\_test} with this bug prints the following transcript output:
\newpage
\begin{minted}[frame=single,framesep=10pt,breaklines]{text}
# UVM_INFO @ 0: reporter [RNTST] Running test gcd_random_test...
# [115] DRIVING: a=24 b=18 in_valid=1
# [125] WATCHDOG ARMED a=24 b=18
# [125] TIMEOUT LIMIT = 13 cycles
# [125] INPUT HANDSHAKE COMPLETE
# Waiting for output handshake... out_valid: 0, out_ready: 1
# 
# ** Fatal: [TIMEOUT] DUT hung! Exceeded max theoretical cycles (13)
#    Time: 255 ns  Scope: testbench_gcd.gcd_if_inst File: testbench_gcd.sv Line: 76
#
# ** Error: TIMEOUT: Module exceeded maximum cycles without asserting out_valid!
#    Time: 255 ns  Scope: testbench_gcd.gcd_if_inst.assert_no_infinite_hang File: testbench_gcd.sv Line: 154
#
# ** Note: $fatal : Simulation stopped.
\end{minted}

\subsection{Bug Report Ticket}
If this were discovered during a real verification cycle, it would be logged to the design team as follows, to request a fix:

\begin{tcolorbox}[
  colback=white,
  colframe=gray!60!black,
  title=\textbf{\textsf{[BUG-001] Premature IDLE Transition (Skipping DONE State)}},
  fonttitle=\large\bfseries\sffamily,
  colbacktitle=gray!15,
  coltitle=black,
  boxrule=1pt,
  arc=4pt
]
\sffamily % Switch to sans-serif to mimic a web UI
\renewcommand{\arraystretch}{1.5} % Add some breathing room between rows
\begin{tabular}{@{} p{0.22\textwidth} p{0.72\textwidth} @{}}
    \textbf{Component:} & GCD Module / FSM Logic \\
    \textbf{Severity:} & \textcolor{red!80!black}{\textbf{CRITICAL}} (Complete datapath deadlock) \\
    \textbf{Description:} & The FSM fails to enter the \texttt{DONE} state after mathematical convergence in the \texttt{RUN} state. It illegally jumps to \texttt{IDLE}. \\
    \textbf{Steps to Reproduce:} & Run any transaction where $A \neq B$. \\
    \textbf{Expected Behavior:} & When \texttt{a\_next == b\_next}, \texttt{state} should be assigned to \texttt{DONE}, triggering \texttt{out\_valid = 1}. \\
    \textbf{Actual Behavior:} & \texttt{state} is assigned to \texttt{IDLE}. The output handshake is never initiated, stalling downstream consumers indefinitely. \\
\end{tabular}
\end{tcolorbox}